---
title: "2-CUTandTag-RemoveDup-Picard"
output: html_document
date: "2024-02-16"
author: Josiah Murray
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Begin by creating environmental variables and directories.
```{r Create Enviromental Variables}
proj_name <-"AlisonEGS" #Enter a name for your project's directory

scratch_dir <- file.path("scratch", "g", "srrao", "AlisonEGS", proj_name) #This builds a path for my RCC scratch directory

jobreport_dir <- file.path("job_reports")

figure_dir <- file.path("figures")


data_dir <- file.path("data")
rds_file <- file.path(data_dir, "EGSx_bulk.rds")

if (!(dir.exists(figure_dir))){
  dir.create(figure_dir)
}

if (!(dir.exists(data_dir))){
  dir.create(data_dir)
}

#Next we will use the Sys.setenv() command to make these file paths and variable accessible from the linux command line.
Sys.setenv(PROJ_NAME = proj_name) 
Sys.setenv(SCRATCH_DIR = paste0("/", scratch_dir)) #Accessing the scratch directory requires a backslash before the filepath
Sys.setenv(JOBREP_DIR = paste0("/", scratch_dir, "/", jobreport_dir))
```


## Split bam files into hg38 and spike ins
It is useful to remove PCR duplicates from CUT&Tag data because even thought it's possible that some duplicate fragments are real... its difficult to say for sure. We can use picard to mark duplicate reads in the aligned bam files.
```{cat Create Picard Script, engine.opts = list(file = 'scripts/4-sort_bam.sh')}
#!/bin/bash

#SBATCH -p normal # partition name
#SBATCH --job-name=sort_bam # Job name
#SBATCH --ntasks=1 # number of tasks
#SBATCH --time 1-16:00 # time limit (day-hour:min)
#SBATCH --cpus-per-task=25 # number of threads
#SBATCH --mem-per-cpu=7Gb # requested memory
#SBATCH --account=srrao # PI's net ID
#SBATCH -o job_reports/%x.out # File to which standard outpout will be written
#SBATCH -e job_reports/%x.err # File to whcih standard error wil be written
#SBATCH --mail-type=ALL # What email updates to send
#SBATCH --mail-user=alimeyer@mcw.edu # Email adress to send to
echo Starting at $(date)
echo Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}
echo I have ${SLURM_CPUS_ON_NODE} CPUs on compute node $(hostname -s)

# Load required modules
module load samtools
module load picard
module load parallel

#Set directory paths
projPath="/scratch/g/srrao/AlisonEGS"
bamPath="${projPath}/alignment/bam"


##Use picard to mark duplicate reads in the aligned BAM files.
picardCMD="java -jar $PICARD"


# Function to sort BAM files using Picard
sort_bam() {
    local i=$1
    $picardCMD SortSam I="$i".bowtie2.bam O="$i".bowtie2.sorted.bam SORT_ORDER=coordinate
    samtools index "$i".bowtie2.sorted.bam
}


# Export functions and variables for parallel
export -f sort_bam
export picardCMD PICARD

# Change to BAM directory for processing
cd ${bamPath}

# Run sorting in parallel (using 5 jobs at once with 5 cores each = 25 total cores)
ls *.hg38.bowtie2.bam | cut -d . -f 1,2 | sort | uniq | parallel -j 5 sort_bam {}



```

Submit the Sort Bam job to SLURM
```{bash Run sort_bam, engine.opts='-l'}
sbatch scripts/4-sort_bam.sh
```


## Remove PCR Duplicates
```{cat Create script to remove dups from BAM files, engine.opts = list(file = 'scripts/5-picard.sh')}
#!/bin/bash

#SBATCH -p normal # partition name
#SBATCH --job-name=picard # Job name
#SBATCH --ntasks=1 # number of tasks
#SBATCH --time 1-12:00 # time limit (day-hour:min)
#SBATCH --cpus-per-task=24 # number of threads
#SBATCH --mem-per-cpu=7Gb # requested memory
#SBATCH --account=srrao # PI's net ID
#SBATCH --output=%x.out # File to which standard output will be written
#SBATCH -o job_reports/%x.out # File to which standard outpout will be written
#SBATCH -e job_reports/%x.err # File to whcih standard error wil be written
#SBATCH --mail-type=ALL # What email updates to send
#SBATCH --mail-user=alimeyer@mcw.edu # Email adress to send to

echo Starting at $(date)
echo Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}
echo I have ${SLURM_CPUS_ON_NODE} CPUs on compute node $(hostname -s)

projPath="/scratch/g/srrao/AlisonEGS"
bamPath="${projPath}/alignment/bam"
outPath="${bamPath}/picard_summary"

mkdir -p ${outPath}

module load samtools
module load picard
module load parallel

# Store the call for 
picardCMD="java -jar $PICARD"

# Define the processing function
process_bam() {
    local input_bam=$1
    local output_path=$2
    
    # Get base name without extension
    local base=$(basename "$input_bam" .bowtie2.bam)
    
    samtools index "$input_bam"
    
    $picardCMD MarkDuplicates \
        I="$input_bam" \
        O="${base}.bowtie2.rmDup.bam" \
        TAG_DUPLICATE_SET_MEMBERS=true \
        REMOVE_DUPLICATES=true \
        METRICS_FILE="${output_path}/${base}.picard.rmDup.txt"
    
    samtools index "${base}.bowtie2.rmDup.bam"
}

export -f process_bam
export picardCMD
export outPath

# Process hg38 BAMs. Picard can use up to 6 threads.
cd $bamPath
echo "Processing hg38 BAMs in ${hg38Path}"
ls *.bowtie2.sorted.bam | parallel -j 4 process_bam {} $outPath

echo "Job completed at $(date)"

```


```{bash Run Bam Split, engine.opts='-l'}
sbatch scripts/5-picard.sh
```



```{cat Create Script for Counting Mito Reads, engine.opts = list(file = 'scripts/6-mito.sh')}
#!/bin/bash

#SBATCH -p normal # partition name
#SBATCH --job-name=mito # Job name
#SBATCH --ntasks=1 # number of tasks
#SBATCH --time 0-02:00 # time limit (day-hour:min)
#SBATCH --cpus-per-task=12 # number of threads
#SBATCH --mem-per-cpu=7Gb # requested memory
#SBATCH --account=srrao # PI's net ID
#SBATCH --output=%x.out # File to which standard output will be written
#SBATCH -o job_reports/%x.out # File to which standard outpout will be written
#SBATCH -e job_reports/%x.err # File to whcih standard error wil be written
#SBATCH --mail-type=ALL # What email updates to send
#SBATCH --mail-user=alimeyer@mcw.edu # Email adress to send to

echo Starting at $(date)
echo Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}
echo I have ${SLURM_CPUS_ON_NODE} CPUs on compute node $(hostname -s)

projPath="/scratch/g/srrao/AlisonEGS"
bamPath="${projPath}/alignment/bam/hg38bams"
outPath="${bamPath}/mito_summary"

mkdir $outPath

cd $bamPath

module load samtools

for i in $(ls *.hg38.bowtie2.bam | cut -d . -f 1 | sort | uniq)
do

#Use samtools to count the number of mitochondrial alignments
mitoDepthDouble=$(samtools view -@ 12 -c "$i".hg38.bowtie2.bam MT)
mitoDepth=$((mitoDepthDouble/2))

echo $mitoDepth >$outPath/"$i".mitoDepth

done
```


```{bash Run Mito, engine.opts='-l'}
sbatch scripts/6-mito.sh
```


## Create Sample Lists and Reading Data into R

Before we create sample lists let's load the necessary R packages.
```{r Load R Packages, echo=FALSE}
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(viridis)
library(corrplot)
library(GenomicRanges)
library(Rsamtools)
```


Read the sample list text files into R
```{r Read Sample Lists into R}
bulk_sampleList <- scan(file = "BulkSampleList.txt", what = "character")
```


Now we have an object in R that is a list of all the sample names for either bulk or single cell CUT&Tag. Next lets enter in the different histone marks that we used in this assay.
```{r Make a Histone Mark List}
bulk_histList <- c("CTCF", "IgG")
```

```{r Read Alignment Summary RDS}
alignSummary <- readRDS(file = "alignSummary.rds")
```

##Summarize Duplicate Information from Picard

We now have the sequencing depth and alignment rates for each of our samples... but how many of the paired end reads are duplicates? We can find this out by looking at the Picard reports.
```{r Summarize Picard Duplicates for Bulk Samples}
bulk_dupResult = c()
for(hist in bulk_sampleList){
  
  
  
  dupRes = read.table(paste0("alignment/bam/picard_summary/", hist, ".hg38.bowtie2.sorted.bam.picard.rmDup.txt"), header = TRUE, fill = TRUE)
  
  histInfo = strsplit(hist, "_")[[1]]
  
  bulk_dupResult = data.frame(Genotype = paste0(histInfo[1]), 
                              Replicate = histInfo[3],
                              Target = histInfo[2],
                              Hg38_MappedFragNum = dupRes$READ_PAIRS_EXAMINED[1] %>% as.character %>% as.numeric, 
                              Hg38_DuplicationRate = dupRes$PERCENT_DUPLICATION[1] %>% as.character %>% as.numeric * 100,
                              Hg38_EstimatedLibrarySize = dupRes$ESTIMATED_LIBRARY_SIZE[1] %>% as.character %>% as.numeric) %>%
                              mutate(Hg38_UniqueMappedFragNum = Hg38_MappedFragNum * (1-Hg38_DuplicationRate/100)) %>%
                                      rbind(bulk_dupResult, .)
}

bulk_dupResult$Genotype = factor(bulk_dupResult$Genotype)

bulk_alignDupSummary = left_join(alignSummary, bulk_dupResult, by = c("Genotype", "Replicate", "Target"))

bulk_alignDupSummary
```


```{r differntiate between clone and genotype in df}


```



```{r}
# Save the bulk_alignDupSummary dataframes as an rds object.
if(!file.exists("alignSummary.rds")){
saveRDS(bulk_alignDupSummary,
        file = "alignSummary.rds")
}

# Read the RDS file 
bulk_alignDupSummary <- readRDS("alignSummary.rds")
```


```{r Graph bulk alignment & duplication}
dupFig1a <- bulk_alignDupSummary %>%
  filter(grepl("IgG", Target)) %>%
  ggplot(aes(x = Genotype, y = Hg38_DuplicationRate)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color = Replicate), position = position_jitter(0.15)) +
  ylab("Duplication Rate (%)") +
  xlab("") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 9), plot.title = element_text(size = 18, hjust = 0.5, margin = margin(0,0,10,0))) +
  guides(fill="none") +
  ggtitle("AlisonEGS - Hg38 Duplication Rate")


dupFig1b <- bulk_alignDupSummary %>%
  filter(grepl("CTCF", Target)) %>%
  ggplot(aes(x = Genotype, y = Hg38_DuplicationRate)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color = Replicate), position = position_jitter(0.15)) +
  ylab("Duplication Rate (%)") +
  xlab("") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 9), plot.title = element_text(size = 18, hjust = 0.5, margin = margin(0,0,10,0))) +
  guides(fill="none") +
  ggtitle("AlisonEGS - Hg38 Duplication Rate") 

dupFig1a
dupFig1b
```


```{r Save Bulk Duplication Rate Figure}
ggsave(filename = "AlisonEGS_DupRate-Hg38_Igg.pdf", 
       plot = dupFig1a, 
       width = 8, 
       height = 6,
       units = "in",
       dpi = "print",
       bg = "white",
       path = "figures")

ggsave(filename = "AlisonEGS_DupRate-Hg38_CTCF.pdf", 
       plot = dupFig1b, 
       width = 8, 
       height = 6,
       units = "in",
       dpi = "print",
       bg = "white",
       path = "figures")

```



```{r Graph bulk alignment & duplication}
dupFig2a <- bulk_alignDupSummary %>%
  filter(grepl("IgG", Target)) %>%
  ggplot(aes(x = Genotype, y = SpikeIn_DuplicationRate)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color = Replicate), position = position_jitter(0.15)) +
  ylab("Duplication Rate (%)") +
  xlab("") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 9), plot.title = element_text(size = 18, hjust = 0.5, margin = margin(0,0,10,0))) +
  guides(fill="none") +
  ggtitle("AlisonEGS - Murine Duplication Rate") +
  facet_wrap(~ Target)

dupFig2b <- bulk_alignDupSummary %>%
  filter(!grepl("CTCF", Target)) %>%
  ggplot(aes(x = Genotype, y = SpikeIn_DuplicationRate)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color = Replicate), position = position_jitter(0.15)) +
  ylab("Duplication Rate (%)") +
  xlab("") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 9), plot.title = element_text(size = 18, hjust = 0.5, margin = margin(0,0,10,0))) +
  guides(fill="none") +
  ggtitle("AlisonEGS - Murine Duplication Rate") +
  facet_wrap(~ Target) 

```

## Aligned Fragments

```{r Plot Aligned Frags}
alignfragFig1a <- bulk_alignDupSummary %>%
  filter(grepl("IgG", Target)) %>%
  ggplot(aes(x = Genotype, y = Hg38_MappedFragNum)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color = Replicate), position = position_jitter(0.15)) +
  ylab("Aligned Fragments") +
  xlab("") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 9), plot.title = element_text(size = 18, hjust = 0.5, margin = margin(0,0,10,0))) +
  guides(fill="none") +
  ggtitle("AlisonEGS - Hg38 Aligned Frags") +
  facet_wrap(~ Target)

alignfragFig1b <- bulk_alignDupSummary %>%
  filter(grepl("CTCF", Target)) %>%
  ggplot(aes(x = Genotype, y = Hg38_MappedFragNum)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color = Replicate), position = position_jitter(0.15)) +
  ylab("Aligned Fragments") +
  xlab("") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 9), plot.title = element_text(size = 18, hjust = 0.5, margin = margin(0,0,10,0))) +
  guides(fill="none") +
  ggtitle("AlisonEGS - Hg38 Aligned Frags") +
  facet_wrap(~ Target)


alignfragFig1a
alignfragFig1b
```


##Estimated Library Sizes

```{r Plot Estimated Library Sizes}
libsizeFig1a <- bulk_alignDupSummary %>%
  #filter(grepl("IgG", Target)) %>%
  ggplot(aes(x = Genotype, y = Hg38_EstimatedLibrarySize)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color = Replicate), position = position_jitter(0.15)) +
  ylab("Estimated Library Size") +
  xlab("") +
  theme_classic() +
  scale_y_continuous(trans='log10') +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 9), 
        plot.title = element_text(size = 18, hjust = 0.5, margin = margin(0,0,10,0))) +
  guides(fill="none") +
  ggtitle("AlisonEGS - Est Library Size") +
  facet_wrap(~ Target)


libsizeFig1b <- bulk_alignDupSummary %>%
  filter(grepl("CTCF", Target)) %>%
  ggplot(aes(x = Genotype, y = Hg38_EstimatedLibrarySize)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color = Replicate), position = position_jitter(0.15)) +
  ylab("Estimated Library Size") +
  xlab("") +
  theme_classic() +
  scale_y_continuous(trans='log10') +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 9), 
        plot.title = element_text(size = 18, hjust = 0.5, margin = margin(0,0,10,0))) +
  guides(fill="none") +
  ggtitle("AlisonEGS - Est Library Size") +
  facet_wrap(~ Target)


libsizeFig1a
libsizeFig1b

```






```{r Save Estimated Library Size Figure}
ggsave(filename = "AlisonEGS_EstLibSize-Hg38_IgG.pdf", 
       plot = libsizeFig1a, 
       device = "pdf",
       width = 8, 
       height = 6,
       units = "in",
       dpi = "print",
       bg = "white",
       path = "figures")

ggsave(filename = "AlisonEGS_EstLibSize-Hg38_CTCF.pdf", 
       plot = libsizeFig1b, 
       device = "pdf",
       width = 8, 
       height = 6,
       units = "in",
       dpi = "print",
       bg = "white",
       path = "figures")
```



## View Spike-In Info

```{r Plot Spike-In Read Percentage}
spikePercFig1a <- bulk_alignDupSummary %>%
  filter(grepl("H3", Target)) %>%
  ggplot(aes(x=Genotype, y=(SpikeIn_MappedFragNum /(Hg38_MappedFragNum + SpikeIn_MappedFragNum)) * 100)) +
  geom_boxplot(outlier.shape = NA) +
  ylab("Percentage of Reads from Mus Musculus") +
  xlab("") +
  theme_classic() +
  geom_jitter(aes(color = Replicate), position = position_jitter(0.15)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 9), plot.title = element_text(size = 12, hjust = 0.5, margin = margin(0,0,10,0))) +
  ggtitle("Murine Spike-In: Read Percentage") +
  facet_wrap(~ Target)


spikePercFig1b <- bulk_alignDupSummary %>%
  filter(grepl("H3", Target)) %>%
  ggplot(aes(x=Genotype,y=(SpikeIn_UniqueMappedFragNum /(Hg38_UniqueMappedFragNum + SpikeIn_UniqueMappedFragNum)) * 100)) +
  geom_boxplot(outlier.shape = NA) +
  ylab("Percentage of De-duplicated Reads from Mus Musculus") +
  xlab("") +
  theme_classic() +
  geom_jitter(aes(color = Replicate), position = position_jitter(0.15)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 9), plot.title = element_text(size = 12, hjust = 0.5, margin = margin(0,0,10,0))) +
  ggtitle("Murine Spike-In: De-duplicated Read Percentage") +
  facet_wrap(~ Target)


spikePercFig1a
spikePercFig1b
```

```{r}
spikePercFig2a <- bulk_alignDupSummary %>%
  filter(!grepl("H3", Target)) %>%
  ggplot(aes(x=Genotype, y=(SpikeIn_MappedFragNum /(Hg38_MappedFragNum + SpikeIn_MappedFragNum)) * 100)) +
  geom_boxplot(outlier.shape = NA) +
  ylab("Percentage of Reads from Mus musculus") +
  xlab("") +
  ylim(c(0,80)) +
  theme_classic() +
  geom_jitter(aes(color = Replicate), position = position_jitter(0.15)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 9), plot.title = element_text(size = 12, hjust = 0.5, margin = margin(0,0,10,0))) +
  ggtitle("Percent of Spike-In Reads: Murine") +
  facet_wrap(~ Target)

spikePercFig2b <- bulk_alignDupSummary %>%
  filter(!grepl("H3", Target)) %>%
  ggplot(aes(x=Genotype, y=(SpikeIn_UniqueMappedFragNum /(Hg38_UniqueMappedFragNum + SpikeIn_UniqueMappedFragNum)) * 100)) +
  geom_boxplot(outlier.shape = NA) +
  ylab("Percentage of De-duplicated Reads from Mus musculus") +
  xlab("") + 
  ylim(c(0,80)) +
  theme_classic() +
  geom_jitter(aes(color = Replicate), position = position_jitter(0.15)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 9), plot.title = element_text(size = 12, hjust = 0.5, margin = margin(0,0,10,0))) +
  ggtitle("Percent of De-duplicated Spike-In Reads: Murine") +
  facet_wrap(~ Target)

spikePercFig2a
spikePercFig2b
```



## Explore correlations between Spike-In % and Other factors

```{r Corr b/w replicate and % Spike In}
bulk_alignDupSummary %>%
  filter(SpikeIn == "murine") %>%
  ggplot(aes(x=Replicate, y=(SpikeIn_MappedFragNum /(Hg38_MappedFragNum + SpikeIn_MappedFragNum)) * 100)) +
  geom_boxplot(outlier.shape = NA) +
  ylab("Percentage of Reads from Mus Musculus") +
  xlab("Replicate") +
  theme_classic() +
  geom_jitter(aes(color = Target), position = position_jitter(0.15)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 9), plot.title = element_text(size = 12, hjust = 0.5, margin = margin(0,0,10,0))) +
  ggtitle("Spike-In % by Replicate")
```


```{r Corr b/w read depth and % Spike In}
bulk_alignDupSummary %>%
  filter(SpikeIn == "murine") %>%
  ggplot(aes(x=SequencingDepth, 
             y=(SpikeIn_MappedFragNum /(Hg38_MappedFragNum + SpikeIn_MappedFragNum)) * 100,
             colour = Target)) +
  geom_point() +
  ylab("Percentage of Reads from Mus Musculus") +
  xlab("Sequencing Depth") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 9), plot.title = element_text(size = 12, hjust = 0.5, margin = margin(0,0,10,0))) +
  ggtitle("Read Depth vs Spike-In %")
```

This next plot clearly shows that increased spike-in correlates with decreases alignment rate... probably because it increases the number of repetitive regions which can lead to multi and or discordant mappings (which are rejected).
```{r Corr b/w % Align and % Spike In}
bulk_alignDupSummary %>%
  filter(SpikeIn == "murine") %>%
  ggplot(aes(x=AlignmentRate, y=(SpikeIn_MappedFragNum /(Hg38_MappedFragNum + SpikeIn_MappedFragNum)) * 100,
             colour = Target)) +
  geom_point() +
  ylab("Percentage of Reads from Mus Musculus") +
  xlab("Alignment Rate") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 9), plot.title = element_text(size = 12, hjust = 0.5, margin = margin(0,0,10,0))) +
  ggtitle("Spike-In % by Replicate")


bulk_alignDupSummary %>%
  filter(Target == "RAD21") %>%
  ggplot(aes(x=AlignmentRate, y=(SpikeIn_MappedFragNum /(Hg38_MappedFragNum + SpikeIn_MappedFragNum)) * 100,
             colour = Replicate)) +
  geom_point() +
  ylab("Percentage of Reads from Mus Musculus") +
  xlab("Alignment Rate") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 9), plot.title = element_text(size = 12, hjust = 0.5, margin = margin(0,0,10,0))) +
  ggtitle("Spike-In % by Replicate")
```


```{r Corr b/w hg38 reads and % Spike In}
bulk_alignDupSummary %>%
  filter(SpikeIn == "murine") %>%
  ggplot(aes(x=Hg38_MappedFragNum, 
             y=(SpikeIn_MappedFragNum /(Hg38_MappedFragNum + SpikeIn_MappedFragNum)) * 100,
             colour = Target)) +
  geom_point() +
  ylab("Percentage of Reads from Mus Musculus") +
  xlab("Hg38 Mapped Frags") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 9), plot.title = element_text(size = 12, hjust = 0.5, margin = margin(0,0,10,0))) +
  ggtitle("Read Depth vs Spike-In %")
```


```{r Corr b/w hg38 reads and % Spike In}
bulk_alignDupSummary %>%
  filter(SpikeIn == "murine") %>%
  ggplot(aes(x=SpikeIn_MappedFragNum, 
             y=(SpikeIn_MappedFragNum /(Hg38_MappedFragNum + SpikeIn_MappedFragNum)) * 100,
             colour = Target)) +
  geom_point() +
  ylab("Percentage of Reads from Mus Musculus") +
  xlab("Spike-In Mapped Frags") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 9), plot.title = element_text(size = 12, hjust = 0.5, margin = margin(0,0,10,0))) +
  ggtitle("Read Depth vs Spike-In %")
```



## Calculate ChIP-Rx normalization
```{r Normalize batch-to-batch variation in Spike-Ins}
chiprx_all_df <- bulk_alignDupSummary %>%
  # Get the chiprx norm factor: 1 over # of spike in reads (in millions)
  mutate(chiprx_fac =  1/(SpikeIn_MappedFragNum/1e6),
         SpikeIn_ReadPercentage = SpikeIn_MappedFragNum/(Hg38_MappedFragNum + SpikeIn_MappedFragNum)) %>%
  group_by(Replicate, Target) %>%
  # Adjust the chiprx norm factor based on the average chiprx norm factor for each replicate
  mutate(chiprx_repMean = mean(chiprx_fac),
         repMean_MappedFragNum = mean(Hg38_MappedFragNum)) %>% ungroup() %>%
  # Calculate a new chiprx norm factor based on each replicate mean
  mutate(chiprx_norm_fac = chiprx_fac / chiprx_repMean) %>%
  # Calculate normalized library sizes
  mutate(chiprx_lib_size = chiprx_fac * Hg38_MappedFragNum,
         chiprx_norm_lib_size = chiprx_norm_fac * Hg38_MappedFragNum) %>%
  group_by(Target) %>%
  mutate(chiprx_sd = sd(chiprx_fac),
         chiprx_norm_sd = sd(chiprx_norm_fac))



chiprx_dupRm_df <- bulk_alignDupSummary %>%
  # Get the chiprx norm factor
  mutate(chiprx_fac =  1/(SpikeIn_UniqueMappedFragNum/1e6),
         SpikeIn_ReadPercentage = SpikeIn_MappedFragNum/(Hg38_MappedFragNum + SpikeIn_MappedFragNum)) %>%
  group_by(Replicate, Target) %>%
  # Adjust the chiprx norm factor based on the average chiprx norm factor for each replicate
  mutate(chiprx_repMean = mean(chiprx_fac)) %>% ungroup() %>%
  # Calculate a new chiprx norm factor based on each replicate mean
  mutate(chiprx_norm_fac = chiprx_fac / chiprx_repMean) %>%
  # Calculate normalized library sizes
  mutate(chiprx_lib_size = chiprx_fac * Hg38_UniqueMappedFragNum,
         chiprx_norm_lib_size = chiprx_norm_fac * Hg38_UniqueMappedFragNum) %>%
  group_by(Target) %>%
  mutate(chiprx_sd = sd(chiprx_fac),
         chiprx_norm_sd = sd(chiprx_norm_fac))
```


```{r}
rx_fig_1a <- chiprx_all_df %>%
  ggplot(aes(x=Genotype, y=chiprx_norm_lib_size)) +
  geom_boxplot(outlier.shape = NA) +
  ylab("Normalized Library Size (RRPM)") +
  xlab("") +
  geom_jitter(aes(color = Replicate), position = position_jitter(0.15)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 9), 
        plot.title = element_text(size = 12, hjust = 0.5, margin = margin(0,0,10,0))) +
  ggtitle("ChIP-Rx Normalized Lib Size: Raw - Within Rep Norm") + 
  facet_wrap(~ Target)


rx_fig_1b <- chiprx_dupRm_df %>%
  ggplot(aes(x=Genotype, y=chiprx_norm_lib_size)) +
  geom_boxplot(outlier.shape = NA) +
  ylab("Normalized Library Size (RRPM)") +
  xlab("") +
  geom_jitter(aes(color = Replicate), position = position_jitter(0.15)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 9), 
        plot.title = element_text(size = 12, hjust = 0.5, margin = margin(0,0,10,0))) +
  ggtitle("ChIP-Rx Normalized Lib Size: Raw - Within Rep Norm") + 
  facet_wrap(~ Target)


rx_fig_1a

rx_fig_1b
```

```{r t tests}
rad21h_test <- chiprx_all_df %>%
  filter(Target == "RAD21" & (Genotype == "RAD21h" | Genotype == "NTC")) %>%
  t.test(chiprx_norm_lib_size ~ Genotype, data = .)

smc3h_test <- chiprx_all_df %>%
  filter(Target == "RAD21" & (Genotype == "SMC3h" | Genotype == "NTC")) %>%
  t.test(chiprx_norm_lib_size ~ Genotype, data = .)

stag2n_test <- chiprx_all_df %>%
  filter(Target == "RAD21" & (Genotype == "STAG2n" | Genotype == "NTC")) %>%
  t.test(chiprx_norm_lib_size ~ Genotype, data = .)


igg_rad21h_test <- chiprx_all_df %>%
  filter(Target == "IgG" & (Genotype == "RAD21h" | Genotype == "NTC")) %>%
  t.test(chiprx_norm_lib_size ~ Genotype, data = .)

igg_smc3h_test <- chiprx_all_df %>%
  filter(Target == "IgG" & (Genotype == "SMC3h" | Genotype == "NTC")) %>%
  t.test(chiprx_norm_lib_size ~ Genotype, data = .)

igg_stag2n_test <- chiprx_all_df %>%
  filter(Target == "IgG" & (Genotype == "STAG2n" | Genotype == "NTC")) %>%
  t.test(chiprx_norm_lib_size ~ Genotype, data = .)

format_pvalue <- function(p) {
  if(p < 0.001) return("p < 0.001")
  else return(paste("p =", round(p, 3)))
}
```


```{r}
# Plot for rad21 chip-rx normalized lib size
fig_rad21_chiprx <- chiprx_all_df %>%
  filter(Target == "RAD21") %>%
  ggplot(aes(x=Genotype, y=chiprx_norm_lib_size)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color = Replicate), position = position_jitter(0.15)) +
  ylab("Normalized Library Size (RRPM)") +
  xlab("") +
  # Add comparison for STAG2n
  annotate("segment", x = 1, xend = 2, y = 25e6, yend = 25e6) +
  annotate("text", x=1.5, y=26e6, label = format_pvalue(stag2n_test$p.value)) +
  # Add comparison for RAD21h
  annotate("segment", x = 1, xend = 3, y = 27.5e6, yend = 27.5e6) +
  annotate("text", x=2, y=28.5e6, label = format_pvalue(rad21h_test$p.value)) +
  # Add comparison for SMC3h
  annotate("segment", x = 1, xend = 4, y = 30e6, yend = 30e6) +
  annotate("text", x=2.5, y=31e6, label = format_pvalue(smc3h_test$p.value)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 9),
        plot.title = element_text(size = 12, hjust = 0.5, margin = margin(0,0,10,0))) +
  ggtitle("Spike-In Normalized CUT&Tag - RAD21 Library Size")


# Plot for igg chip-rx normalized lib size
fig_igg_chiprx <- chiprx_all_df %>%
  filter(Target == "IgG") %>%
  ggplot(aes(x=Genotype, y=chiprx_norm_lib_size)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color = Replicate), position = position_jitter(0.15)) +
  ylab("Normalized Library Size (RRPM)") +
  xlab("")  +
  # Add comparison for STAG2n
  annotate("segment", x = 1, xend = 2, y = 13e6, yend = 13e6) +
  annotate("text", x=1.5, y=13.6e6, label = format_pvalue(igg_stag2n_test$p.value)) +
  # Add comparison for RAD21h
  annotate("segment", x = 1, xend = 3, y = 14.4e6, yend = 14.4e6) +
  annotate("text", x=2, y=15e6, label = format_pvalue(igg_rad21h_test$p.value)) +
  # Add comparison for SMC3h
  annotate("segment", x = 1, xend = 4, y = 15.8e6, yend = 15.8e6) +
  annotate("text", x=2.5, y=16.4e6, label = format_pvalue(igg_smc3h_test$p.value)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 9),
        plot.title = element_text(size = 12, hjust = 0.5, margin = margin(0,0,10,0))) +
  ggtitle("Spike-In Normalized CUT&Tag - IgG Library Size")

fig_rad21_chiprx
fig_igg_chiprx
```



```{r Save Spike-In Figures}
ggsave(filename = "OCIAML3_CnT-Rx_LibSize_RAD21_Spike-In-Norm.pdf", 
       plot = fig_rad21_chiprx, 
       device = "pdf",
       width = 5, 
       height = 4,
       units = "in",
       dpi = "print",
       bg = "white",
       path = "figures")

ggsave(filename = "OCIAML3_CnT-Rx_LibSize_IgG_Spike-In-Norm.pdf", 
       plot = fig_igg_chiprx, 
       device = "pdf",
       width = 5, 
       height = 4,
       units = "in",
       dpi = "print",
       bg = "white",
       path = "figures")
```


```{r}
# export chiprx normalization factors to a csv
chiprx_all_df %>%
  ungroup() %>%
  mutate(Sample = paste(Clone, Target, Replicate, sep = "_")) %>%
  select(Sample, chiprx_norm_fac) %>%
  write_delim(file = "normalization_factors.tsv", delim = "\t")


# add chipr rx data to alignment summary
saveRDS(chiprx_all_df,
        file = "chiprx.rds")

chiprx_all_df <- readRDS(file = "chiprx.rds")
```


##Print session info
```{r session info}
sessionInfo()
```

